\renewcommand{\thepage}{Biographical Sketch  - Page \arabic{page} of 2}
\required{Biographical Sketch --- Kimberly Julie Gilbert}

\begin{comment}
Biographical Sketch [2 page limit]. Submit your CV in the section Biographical Sketches. Follow the
form in the Grants Program Guide II.C.2.f.i for senior personnel. List conference abstracts and titles of
presentations separately from peer--?reviewed articles. For unpublished manuscripts, list only those
submitted or accepted for publication (along with most likely date of publication).
\end{comment}

% Your Bio should be divided into the following sections

%	I.	Professional Preparation
%		------------------------------------------

%	*	Mention a complete list of undergraduate and graduate education and postdoctoral 		                     training as shown:

%		Institution(s)			Major			Degree		Year
%		------------------- 			--------- 			------------ 		-------
%		Undergraduate
%		Graduate
%		Postdoctoral			(Area)			N/A
%		-------------------------------------------------------------------------------------------------------

\setcounter{page}{1}
\renewcommand{\thepage}{Biographical Sketch - Page \arabic{page} of 2}
%\begin{center}
%\textbf{\large Biographical Sketch}
%\end{center}


\textbf{A.  Professional Preparation}
\vspace{2pt}	

\begin{tabular}{llll}
Institution, Location \hspace{0.5in} & Major \hspace{1in} & Degree  \hspace{0.25in} & Year \\
\hline
University of Virginia, USA  & Biology  & B.Sc.   & 2010 \\
University of British Columbia, Canada                     & Zoology              & Ph.D.                  & 2016 (expected) \\
%\hspace{1cm} \emph{Advisor}: Michael C. Whitlock
\end{tabular}

%	II.	Appointments
%		------------------------

%	*	A list in reverse chronological order of all academic / professional appointments
%		beginning with the current appointment as shown:

%		------------------------------------------------------------------------
%		Year(s)	Position, Department, Institution
%		------------------------------------------------------------------------

%\begin{tabular}{ll}
%Year-present & Position, Department, Institution \\
%Year(s)      & Position, Department, Institution \\
%\end{tabular}


%	III. 	Publications
%		---------------------

%	*	Up to 5 publicationa most closely related with the porposed project
%	*	Up to 5 other publications, whether or not related to the proposed project
%	* 	Each publication must include the names of all authors in the same sequence as they
%		appear in the publication, the article and journal title, book title, volume number, page
%		numbers and year of publication.
%	*	The website address should be included if the document is available electronically
%	*	List any unpublished manuscripts that have been accepted for publication; also include
%		the most likely date of publication.
%	* 	Patents, copyrights and software systems may be substituted for publications.
%	*	Only the list of 10 publications will be used in the review of the proposal.

\vspace{6pt}
\noindent
\textbf{B.  Publications}
\vspace{2pt}	

\emph{Five Publications Most Closely Related to the Proposed Project}
\vspace{4pt}
%
\kjg{maybe still want to change around which pubs go in 5 most important?}
%
\begin{enumerate}[nolistsep]
\item \textbf{Gilbert KJ}, MC Whitlock (2015) \href{http://onlinelibrary.wiley.com/doi/10.1111/evo.12713/abstract}{Evaluating methods for estimating local effective population size with and without migration}. \emph{Evolution}, 68(8), 2154-2166.


\item \textbf{Gilbert KJ}, MC Whitlock (2015) \href{http://onlinelibrary.wiley.com/doi/10.1111/1755-0998.12303/abstract}{\emph{Q}\footnotesize{\textsc{st}}-\normalsize \emph{F}\footnotesize{\textsc{st}} \normalsize comparisons with unbalanced half-sib designs}. \emph{Molecular Ecology Resources}, 15(2), 262-267.

\item Caplins SA, \textbf{KJ Gilbert}, C Ciotir, J Roland, SF Matter, N Keyghobadi (2014) \href{http://m.rspb.royalsocietypublishing.org/content/281/1796/20141798.abstract?cpetoc}{Landscape structure and the genetic effects of a population collapse}. \emph{Proceedings of the Royal Society B}. 281: 20141798; doi: 10.1098/rspb.2014.1798 
   
\item Keller SR, \textbf{KJ Gilbert}, PD Fields, DR Taylor (2012) \href {http://onlinelibrary.wiley.com/doi/10.1111/j.1365-294X.2012.05751.x/abstract}{Bayesian inference of a complex invasion history revealed by nuclear and chloroplast genetic diversity in the colonizing plant, \emph{Silene latifolia}}. \emph{Molecular Ecology}, 21(19), 4721-4734.

\item Whitlock MC, \textbf{KJ Gilbert} (2012) \href {http://onlinelibrary.wiley.com/doi/10.1111/j.1755-0998.2012.03122.x/abstract}{ \emph{Q}\footnotesize{\textsc{st}} \normalsize{in a hierarchically structured population}. \emph{Molecular Ecology Resources}, 12(3), 481--483.} 
    
\end{enumerate}

\vspace{2pt}
\noindent
~ \emph{Four Other Significant Publications}
\vspace{2pt}

\begin{enumerate}[nolistsep]
   
\item Santiso X, L Lopez, \textbf{KJ Gilbert}, R Barreiro, MC Whitlock, R Retuerto (2015) \href{http://www.sciencedirect.com/science/article/pii/S1433831915000190}{Patterns of genetic variation within and among populations in \emph{Arbutus unedo} and its relation with selection and evolvability.} \emph{Perspectives in Plant Ecology, Evolution and Systematics}, 17(3), 185-192.

\item Vines TH, RL Andrew, DG Bock, MT Franklin, \textbf{KJ Gilbert}, NC Kane, EJ Kleynhans, J-S Moore, BT Moyers, S Renaut, DJ Rennison, T Veen, S Yeaman (2013) \href{http://www.fasebj.org/content/early/2012/12/28/fj.12-218164.abstract?sid=f95019e2-6a54-4da6-b35b-e36826bfb06e}{Mandated archiving greatly improves access to research data}. \emph{FASEB Journal}, 27(4), 1304-1308.  

\item \textbf{Gilbert KJ}, RL Andrew, DG Bock, MT Franklin, NC Kane, J-S Moore, BT Moyers, S Renaut, DJ Rennison, T Veen, TH Vines (2012) \href {http://onlinelibrary.wiley.com/doi/10.1111/j.1365-294X.2012.05754.x/abstract}{Recommendations for utilizing and reporting population genetic analyses: The reproducibility of genetic clustering using the program \textsc{structure}}. \emph{Molecular Ecology}, 21(20), 4925-4930.

\item Vines TH, AYK Albert, RL Andrew, F D\'ebarre, DG Bock, MT Franklin, \textbf{KJ Gilbert}, J-S Moore, S Renaut, DJ Rennison (2014) \href{http://www.sciencedirect.com/science/article/pii/S0960982213014000}{The availability of research data declines rapidly with age}. \emph{Current Biology}, 24, 94-97.

\end{enumerate}

\vspace{4pt}	

\noindent \textbf{C.  Select Conference Presentations} (chosen from 11 presentations)
\vspace{2pt}	
\begin{enumerate}[nolistsep]

\item[2015] Validating SNP loci underlying local adaptation in lodgepole pine; KJ Gilbert, S Yeaman, KE Lotterhos, KA Hodgins, H Suren, JA Holliday, S Nadeau, SN Aitken, MC Whitlock ~ \emph{Poster, 15$^{\textrm{th}}$ ESEB Congress, Lausanne, Switzerland}
	
\item[2014] Evaluating methods for estimating effective population size in the presence of migration; KJ Gilbert \& MC Whitlock ~ \emph{Oral presentation, Evolution, Raleigh, USA}

\item[2012] Range expansion and adaptation across heterogeneous environments; KJ Gilbert \& MC Whitlock ~ \emph{Poster, Evo-WIBO Conference (Evolutionary Biology in the Pacific Northwest), Port Townsend, USA}

\item[2011] Inferred invasion history of \emph{Silene latifolia} into North America utilizing population genetic data and approximate Bayesian computation; KJ Gilbert, SR Keller, PD Fields, DR Taylor ~ \emph{Poster, 13$^{\textrm{th}}$ Congress of the European Society for Evolutionary Biology, Tuebingen, Germany}

\end{enumerate}	
	
\vspace{4pt}	

\noindent
\textbf{D.  Grants and Awards}\\
\vspace{2pt}	
\textmd{Cordula and Gunter Paetzold Fellowship ~\emph{UBC} \$18,000CAD} \hfill {2015 -- 2016}\\
\textmd{\emph{Declined}; Zoology Graduate Fellowship ~\emph{UBC} \$11,000CAD} \hfill {2015 -- 2016}\\
\textmd{Ann and William Messenger Graduate Fellowship ~\emph{UBC} \$700CAD} \hfill {2015}\\
\textmd{Zoology Graduate Fellowship ~\emph{UBC} \$11,000CAD} \hfill {2014 -- 2015}\\
\textmd{Frieda Granot Graduate Scholarship in Interdisciplinary Research ~\emph{UBC} \$200CAD} \hfill {2013 -- 2014}\\
\textmd{Theodore E Arnold Fellowship ~\emph{UBC} \$7,750CAD} \hfill {2013 -- 2014}\\
\textmd{Patrick David Campbell Graduate Fellowship ~\emph{UBC} \$8,050CAD} \hfill {2013 -- 2014}\\
\textmd{\emph{Declined}; Zoology Graduate Fellowship ~\emph{UBC} \$10,000CAD} \hfill {2013 -- 2014}\\
\textmd{BRITE Fellowship ~\emph{UBC} \$10,500CAD \emph{per annum}} \hfill {2011 -- 2013}\\

% goes outside margins
%\begin{tabular}{lll}
%Cordula and Gunter Paetzold Fellowship, UBC \hspace{0.01in} & \$18,000CAD \hspace{0.1in} & 2015 -- 2016 \\
%\emph{Declined}; Zoology Graduate Fellowship, UBC  	&	\$11,000CAD	& 	2015 -- 2016 \\
%Ann and William Messenger Graduate Fellowship, UBC  	&	\$700CAD	& 	2015 \\
%Zoology Graduate Fellowship, UBC  					&	\$11,000CAD 	& 	2014 -- 2015 \\
%Frieda Granot Grad. Scholarship in Interdisciplinary Research & \$200CAD  	& 	2013 -- 2014 \\
%Theodore E Arnold Fellowship  						&	\$7,750CAD 	& 	2013 -- 2014 \\
%Patrick David Campbell Graduate Fellowship  			&	\$8,050CAD 	& 	2013 -- 2014 \\
%\emph{Declined}; Zoology Graduate Fellowship, UBC 	&	\$10,000CAD 	& 	2013 -- 2014 \\
%BRITE Fellowship, UBC  							&	\$21,000CAD 	& 	2011 -- 2013 \\
%\end{tabular}

%	IV.	Synergistic Activities
%		-----------------------------------

%	*	List up to five examples that demonstrate the broader impact of the professional 
%		and scholarly activities that focuses on the integration and knowledge transfer as 
%		well as its creation.
%	*	Examples could
%		include, among others: innovations in teaching and training (e.g., development
%		of curricular materials and pedagogical methods); contributions to the science
%		of learning; development and/or refinement of research tools; computation
%		methodologies, and algorithms for problem-solving; development of databases
%		to support research and education; broadening the participation of groups
%		under-represented in science, mathematics, engineering and technology; and
%		service to the scientific and engineering community outside of the
%		immediate organization.

\noindent \textbf{E.  Synergistic Activities}

\begin{enumerate}[itemsep=-0.25ex]
\item \textbf{Working Groups:} Participated in the \href{http://www.nescent.org/}{NESCent} Reproducible Science Hackathon (2014), a 21-member working group aimed at developing a curriculum and workflow for teaching reproducible science to researchers of any background. Participated in the SimBank \href{http://www.nescent.org/}{NESCent} Catalysis Meeting (2014) which was a 25-member working group to create a collection of openly available simulation results to facilitate testing of statistical population genetic and phylogeographic methods.

\item \textbf{Teaching:} Teaching assistant for Fundamentals of Evolutionary Biology (Fall 2012-Spring 2013) where I taught three sections per term of 45 students each and lead discussion-based tutorials. Teaching assistant for Fundamentals of Biostatistics (Fall 2013, 2014, 2015). Taught two sections of 70 students total in 2013, and in 2014 and 2015 served as the lab coordinator for 254 and 276 students enrolled in the course, respectively, while teaching one section of 36 and 35 students respectively.

\item \textbf{Service:} Served as a Graduate Student Council Member for the \href {http://www.amnat.org/home.html}{American Society of Naturalists} (2013-2016, chair 2015-2016). Served as the graduate student representative on the 2014 evolutionary biology CRC2 job search for the Department of Zoology, University of British Columbia. Organize the Biodiversity Research Centre's weekly \href{http://www.biodiversity.ubc.ca/edg/}{evolution discussion group} (2014-2016), for students, post-docs, and faculty  from the departments of Zoology, Botany, Forestry, and Fisheries to discuss current papers in evolutionary biology. Reviewer for \emph{Molecular Ecology Resources, Ecology and Evolution, Tree Genetics \& Genomes}.

\item \textbf{Outreach:}  Volunteer mist-netting and bird banding with local Vancouver non-profit organization \href {http://www.wildresearch.ca/IonaIsland.html}{Wild Research} (2013-2015) where I participated in winter, spring migration, and fall migration bird monitoring at Iona Island Bird Observatory, taught volunteers proper bird handling, aging, data collection, and mist net extraction techniques, and assisted in educating public visitors to the station about the species conservation and monitoring, and the general tasks of running a banding station.

\end{enumerate}

%%\vspace{12pt}
%%\subsection*{Collaborators \& Other Affiliations}

%%\vspace{12pt}
%%\emph{Collaborators:}
%A list of all persons in alphabetical order (including their current
%organizational affiliations) who are currently, or who have been collaborators
%or co-authors on a project, book, article, report, abstract
%or paper during the 48 months preceding the submission of this proposal.
%Also include those individuals who are currently or have been co-editors
%of a journal, compendium, or conference proceedings during the 24 months
%preceding the submission of the proposal. If there are no collaborators
%or co-editors to report, this should be so indicated.


\vspace{4pt}	

\noindent \textbf{F. Collaborators (Total: 25)}
\vspace{2pt}	

\noindent \emph{U Alberta} Jens Roland; \emph{U Basel} Peter Fields; \emph{U British Columbia} Dan Bock, Diana Rennison; \emph{U Calgary} Sam Yeaman; \emph{UC Davis} Serena Caplins; \emph{U Cincinnati} Stephen Matter; \emph{CIRB Paris} Florence D\'ebarre; \emph{Colorado State} Brook Moyers; \emph{U Colorado} Nolan Kane; \emph{U Coruna} Rodolfo Barreiro, L\'ua L\'opez; \emph{Kwantlen Polytechnic U} Michelle Franklin; \emph{U Laval} Jean-S\'ebastien Moore; \emph{Mol. Ecol. Managing Editor} Timothy Vines; \emph{U Montreal} S\'ebastien Renaut; \emph{U New England (Australia)} Rose Andrew; \emph{UT Austin} Thor Veen; \emph{Trent U} Claudia Ciotir; \emph{U Santiago de Compostela} Rub\'en Retuerto Franco, Xabier Santiso; \emph{U Vermont} Stephen Keller; \emph{U Virginia} Douglas Taylor; \emph{Western U} Nusha Keyghobadi; \emph{Women's Health Research Institute} Arianne Albert\\


%%\vspace{12pt}
%%\emph{Graduate and Postdoctoral Advisors:}
%A list of the names of the individual's own graduate advisor(s) and
%principal postdoctoral sponsor(s), and their current organizational
%affiliations.

\noindent \textbf{Graduate Advisor (Total: 1)} ~ \emph{University of British Columbia} ~ Michael C. Whitlock




%%\vspace{12pt}
%%\emph{Thesis Advisor and Postgraduate-Scholar Sponsor:}
%A list of all persons (including their organizational affiliations),
%with whom you have had an association as thesis advisor, or with
%whom your have had an association within the last five years as a
%postgraduate-scholar sponsor. The total number of graduate students advised
%and postdoctoral scholars sponsored also must be identified.


