\documentclass[10pt]{article}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage{ucs}
\usepackage[utf8x]{inputenc} 
\usepackage{url}
\usepackage[top=2.54cm, bottom=2.54cm, left=2.54cm, right=2.54cm]{geometry}
\usepackage{hyperref}
\usepackage{setspace} 
\newcommand{\ignore}[1]{}
\onehalfspacing
\setstretch{1} 
\usepackage[explicit]{titlesec}
\usepackage[normalem]{ulem} 

\titleformat{\section}{\normalfont}{}{0pt}{\textbf{(\alph{section}) #1}}
%\titleformat{⟨command⟩}[⟨shape⟩]{⟨format⟩}{⟨label⟩}{⟨sep⟩}{⟨before-code⟩}[⟨after-code⟩]
\titleformat{\subsection}
  {\normalfont}{\thesubsection}{1em}{\uline{#1}}
%\usepackage{helvet}
%\renewcommand{\familydefault}{\sfdefault}

%\titleformat{\section}
%  {\normalfont\textbf}{}{}{(\alph{section})\,}
%\renewcommand{\thesection}{(\alph{section})}
%\renewcommand{\thesubsection}{(\uline{subsection})}

\begin{document}

\begin{center}
		\sf\textbf{{Biographical Sketch --- Kevin Richard Thornton}}
\end{center}
%\begin{minipage}{0.55\linewidth}
%  \href{http://www.plantsciences.ucdavis.edu/plantsciences/}Department of Plant Sciences\\
%  \href{http://www.ucdavis.edu/}{University of California} \\
%  1 Shields Ave.\\
%  Davis, CA 95616
%\end{minipage}
%\begin{minipage}{0.35\linewidth}
%  \begin{tabular}{ll}
%    Phone: & (530) 752-1152 \\
%    Fax: &  (530) 752-4604 \\
%    Email: & \href{mailto:rossibarra@ucdavis.edu}{rossibarra@ucdavis.edu} \\
%    Web: & \href{http://www.rilab.org/}{www.rilab.org} \\
%  \end{tabular}
%\end{minipage}
%\bigskip

%\section{Expertise}

%Dr.\ Ross-Ibarra's research focuses on the population genetics of maize and its wild relatives.  His group uses population genetics to investigate the evolution of natural and cultivated populations, identify loci important for adaptation, and understand the evolutionary forces shaping diversity in the maize genome.    
%\bigskip

\section{Professional Preparation}

\begin{tabular}{l l l l}
Institution    \hspace{52mm}              &   Area  \hspace{10mm}     & Degree / Training  \hspace{13mm}    & Dates \\
\hline
University of Puget Sound & Botany & BA& 1997, 2000 \\
University of Chicago & Genetics & PhD & 2003\\
Cornell University & Genetics & Postdoctoral Research & 2007 \\
\end{tabular}

\section{Professional Appointments}

\begin{tabular}{l l l}
Position & Institution                 & Dates\\
\hline
Associate Professor & University of California Irvine &		2012-present \\
Assistant Professor & University of California Irvine &		2007-2012 \\

\end{tabular}

\section{Products}

\subsection*{Most Relevant to the Proposed Research}

\begin{itemize} \setlength{\itemsep}{0pt} \setlength{\parskip}{2pt} \setlength{\parsep}{0pt}

\item \textbf{Thornton}, K. R. (2014) A C++ template library for efficient forward-time population genetic simulation of large populations. Genetics 98:157-166 PMID: 24950894

\item \textbf{Thornton, K}. (2003) libsequence, a C++ class library for evolutionary genetic analysis. Bioinformatics 19(17): 2325-2327 PMID 14630667

\item \textbf{Thornton, K. R.}, A. J. Foran, and A. D. Long (2013) Properties and modeling of GWAS when complex disease risk is due to non-complementing, deleterious mutations in genes of large effect. PLoS Genetics 9: e1003258. PMID 23437004

\end{itemize}

\subsection*{Additional Products}

\begin{itemize} \setlength{\itemsep}{0pt} \setlength{\parskip}{2pt} \setlength{\parsep}{0pt}

\item Cridland, J. M., \textbf{K. R. Thornton} and A. D. Long (2015) Gene expression variation in Drosophila melanogaster due to rare transposable element insertion alleles of large effect. Genetics 199: 85-93.

\item Baldwin-Brown, J., A. D. Long, and \textbf{K. R. Thornton} (2014) The Power to Detect Quantitative Trait Loci Using Resequenced, Experimentally Evolved Populations of Diploid, Sexual Organisms. Molecular Biology and Evolution 31: 1040-1055. PMID 24441104

\item Open-source software: \url{http://molpopgen.github.io/fwdpp/}

\item Open-source software: \url{http://molpopgen.github.io/libsequence/}
\end{itemize}

\section{Synergistic Activities}

\begin{itemize} \setlength{\itemsep}{0pt} \setlength{\parskip}{2pt} \setlength{\parsep}{0pt}
\item Open-source software: \url{http://molpopgen.github.io/fwdpy/}  This software is unpublished, and will be a key resource for this proposal.
\item Editor, G3
\end{itemize}

\section{Collaborators and Other Affiliations}
\subsubsection*{Collaborators and Co-editors (Total: 7)}
\begin{itemize}
\item \emph{Cornell University} Andrew G. Clark
\item \emph{North Carolina State University} Trudy Mackay
\item \emph{Princeton University} Peter Andolfatto 
\item \emph{Rochester University} Daniel Garrigan, Daven C. Presgraves
\item \emph{UC Irvine} Anthony (Tony) Long 
\item \emph{University of Kansas} Stuart MacDonald
\end{itemize}

\subsubsection*{Graduate Advisors and Postdoctoral Sponsors (Total: 2)}

\emph{University of Chicago} Manyuan Long

\emph{Cornell University} Andrew G. Clark

\subsubsection*{Thesis Advisor and Postgraduate Sponsor (Total: 3)}
\textbf{Postdoctoral: } Rebekah R. Rogers
\textbf{PhD Thesis Advisor:} Julie M. Cridland, Jaleal S. Sanjak

\end{document}

